---
title: "The PBTM R-Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{The PBTM R-Package}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
#### Population-based Threshold Models (PBTM) Calculator - beta version. 
This R package combines existing PBTMs calculation, parameters output and respective plots. 

## Installation 
Installation of the package requires the "devtools" package. 

#### Install development version from Github
```
devtools::install_github("pedrobello/PBTM")
```

#### Load package
```
library("PBTM")
```

## Data preparation steps

Before using the PBTM package, data preparation and selection is required. Main steps are described below.

*    Download the template file (download PBTM-R-Data-Empty.csv or the example file PBTM-R-Data-Sample.csv).
*    Enter your data on the proper columns and leave unsed columns in blank.
*    Inform your working folder in R. Add the folder location of your data files.
```
setwd("~/MyProjectFolder/MyDataFolder")
```
*    Load Seeds data - Add the name of the data file (csv file)
```
mydata <- read.csv("PBTM-R-Data-MyData.csv", header=T)
```
*    Filter the data will be used to calculate the model.
```
mydata <- read.csv("PBTM-R-Data-MyData.csv", header=T)
```
*    Reference the filtered data to the RawModelDt data table. 
```
mydata <- read.csv("PBTM-R-Data-MyData.csv", header=T)
```







```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
